Fundamental Problems with the Evidence Base of Adolescent Depression Treatments.

### 2024-02-27 apples_and_oranges.qmd
Age variable transformations

product_perc_women <-  (df_appl_v_orange$percent_women/100)*
  (1-(df_appl_v_orange$percent_women/100) )

total_n <- df_appl_v_orange$baseline_n_active + 
  df_appl_v_orange$baseline_n_control

df_appl_v_orange$percent_women_std_error <- sqrt(product_perc_women/total_n )

@@

# Step 2: turn dataframes to long
# A: turn long the rows with active
active_df <- df_with_distinct_instruments %>% 
  dplyr:: select(new_study_id, active_type, descr_active, psy_or_med, baseline_n_active, cohens_d_active,
                 baseline_mean_active, baseline_sd_active, post_mean_active, post_sd_active,
                 perc_baseline_mean_active, perc_baseline_sd_active, perc_post_mean_active, perc_post_sd_active, percent_women, percent_women_std_error, mean_age, sd_age) %>% 
  rename(treatment = descr_active) %>% 
  rename(type = active_type) %>% 
  rename_with(~ str_remove(., "_active")) %>% 
  mutate(arm_effect_size = "cohens_d_active") %>% # chose this clunky name because I had used this variable name in previous code and don't want to bread the rest
  group_by(new_study_id) %>% #go through each study id and remove duplicates
  distinct(treatment, .keep_all = TRUE)

control_df <- df_with_distinct_instruments %>% 
  dplyr:: select(new_study_id, control_type, descr_control, psy_or_med, baseline_n_control, cohens_d_control,
                 baseline_mean_control, baseline_sd_control,post_mean_control, post_sd_control,
                 perc_baseline_mean_control, perc_baseline_sd_control, perc_post_mean_control, perc_post_sd_control, percent_women, percent_women_std_error, mean_age, sd_age) %>% 
  rename(treatment = descr_control) %>% 
  rename(type = control_type) %>% 
  rename_with(~ str_remove(., "_control")) %>% 
  mutate(arm_effect_size = "cohens_d_control") %>% #chose this clunky name because I had used this variable name in previous code and don't want to bread the rest
  group_by(new_study_id) %>% #go through each study id and remove duplicates
  distinct(treatment, .keep_all = TRUE)

@@

# Turn age sd into se

df_long_for_metan <- df_long_for_metan %>% 
  mutate (se_age = sd_age / sqrt(baseline_n) )

@@

@@

df_for_gender_metan <- list_df_simulated[[1]] %>% 
  filter(!is.na(percent_women))

met_perc_women <- metagen(TE = percent_women,
                           seTE = percent_women_std_error,
                           studlab = new_study_id,
                           data = df_for_gender_metan,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "percentage women across studies",
                           subgroup = four_level_var,
                           tau.common = FALSE)
met_perc_women

@@

met_perc_women_clin <- metagen(TE = percent_women,
                           seTE = percent_women_std_error,
                           studlab = new_study_id,
                           data = df_for_gender_metan_clin,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "percentage women across clinical studies",
                           subgroup = four_level_var,
                           tau.common = FALSE)
met_perc_women_clin

@@

df_for_age_metan <- list_df_simulated[[1]] %>% 
  filter(!is.na(mean_age))

met_age <- metagen(TE = mean_age,
                           seTE = sd_age / sqrt(baseline_n),
                           studlab = new_study_id,
                           data = df_for_age_metan,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "age across studies",
                           subgroup = four_level_var,
                           tau.common = FALSE)

@@

met_age <- metagen(TE = mean_age,
                           seTE = sd_age / sqrt(baseline_n),
                           studlab = new_study_id,
                           data = df_for_age_metan_clin,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "age across clinical studies",
                           subgroup = four_level_var,
                           tau.common = FALSE)

>>>>

@@

# Step 2: turn dataframes to long

# A small bit of cleaning

df_with_distinct_instruments <- df_with_distinct_instruments %>% 
  rename (percent_women_active = active_percent_women, 
          percent_women_control = control_percent_women,
          percent_women_overall = percent_women)

# A: turn long the rows with active
active_df <- df_with_distinct_instruments %>% 
  dplyr:: select(new_study_id, active_type, descr_active, psy_or_med, baseline_n_active, cohens_d_active,
                 baseline_mean_active, baseline_sd_active, post_mean_active, post_sd_active,
                 perc_baseline_mean_active, perc_baseline_sd_active, perc_post_mean_active, perc_post_sd_active, percent_women_overall, percent_women_active, age_m_active, age_sd_active, age_m_overall, age_sd_overall) %>% 
  rename(treatment = descr_active) %>% 
  rename(type = active_type) %>% 
  rename_with(~ str_remove(., "_active")) %>% 
  mutate(arm_effect_size = "cohens_d_active") %>% # chose this clunky name because I had used this variable name in previous code and don't want to bread the rest
  group_by(new_study_id) %>% #go through each study id and remove duplicates
  distinct(treatment, .keep_all = TRUE)

control_df <- df_with_distinct_instruments %>% 
  dplyr:: select(new_study_id, control_type, descr_control, psy_or_med, baseline_n_control, cohens_d_control,
                 baseline_mean_control, baseline_sd_control,post_mean_control, post_sd_control,
                 perc_baseline_mean_control, perc_baseline_sd_control, perc_post_mean_control, perc_post_sd_control, percent_women_overall, percent_women_control, age_m_control, age_sd_control, age_m_overall, age_sd_overall) %>% 
  rename(treatment = descr_control) %>% 
  rename(type = control_type) %>% 
  rename_with(~ str_remove(., "_control")) %>% 
  mutate(arm_effect_size = "cohens_d_control") %>% #chose this clunky name because I had used this variable name in previous code and don't want to bread the rest
  group_by(new_study_id) %>% #go through each study id and remove duplicates
  distinct(treatment, .keep_all = TRUE)

@@

# I'm going to perform  the operations to calculate overall age and percent women here now, I previously did this in the other script, however it actually needs to be done once the dataframe is turned long.

df_long_for_metan <- df_long_for_metan %>% 
  group_by(new_study_id) %>%
  mutate(n_overall = sum(baseline_n)) %>% 
  mutate(age_overall = sum(age_m * baseline_n) / sum(baseline_n)) %>% 
  mutate(mean_age = coalesce(age_overall, age_m_overall)) %>% 
  mutate(pooled_sd_age = sqrt(sum((age_sd^2) * (baseline_n - 1)) / sum(baseline_n - 1))) %>% 
  mutate(se_age = pooled_sd_age / sqrt(n_overall)) %>% 
  mutate(gender_overall = sum(percent_women * baseline_n) / sum(baseline_n)) %>% 
  mutate(percent_women = coalesce(percent_women_overall, gender_overall))

# now just get rid of the unnecessary columns as these are causing clutter

df_long_for_metan <- df_long_for_metan %>% 
  select(-c(percent_women_overall, age_m, age_sd, age_m_overall, age_sd_overall, age_overall, gender_overall))

df_long_for_metan <- df_long_for_metan %>% 
  mutate(product_perc_women = (percent_women/100)* (1-(percent_women/100)) )

df_long_for_metan <- df_long_for_metan %>% 
  mutate(percent_women_std_error = sqrt(product_perc_women/ n_overall ))

@@

# Looking at demographics at baseline
# Currently we have both control and active conditions, we just need each study once

df_baseline_demographics <- df_long_for_metan %>% 
  distinct(new_study_id, .keep_all = TRUE)

@@

df_for_gender_metan <- df_baseline_demographics %>% 
  filter(!is.na(percent_women))

met_perc_women <- metagen(TE = percent_women,
                           seTE = percent_women_std_error,
                           studlab = new_study_id,
                           data = df_for_gender_metan,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "percentage women across studies",
                           subgroup = psy_or_med,
                           tau.common = FALSE)
met_perc_women

@@

met_perc_women_clin <- metagen(TE = percent_women,
                           seTE = percent_women_std_error,
                           studlab = new_study_id,
                           data = df_for_gender_metan_clin,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "percentage women across clinical studies",
                           subgroup = psy_or_med,
                           tau.common = FALSE)
met_perc_women_clin

# Now looking at the CBT + med studies only

df_for_gender_metan_cbt <- df_for_gender_metan %>%
  filter(new_study_id %in% cbt_meds_study_ids) %>% 
  filter(!(arm_effect_size == "cohens_d_active" & psy_or_med == 1 & !(type == "cbt")))

met_perc_women_cbt <- metagen(TE = percent_women,
                           seTE = percent_women_std_error,
                           studlab = new_study_id,
                           data = df_for_gender_metan_cbt,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "percentage women across cbt and med studies",
                           subgroup = psy_or_med,
                           tau.common = FALSE)
met_perc_women_cbt

@@

df_for_age_metan <- df_baseline_demographics %>% 
  filter(!is.na(mean_age))

met_age <- metagen(TE = mean_age,
                           seTE = se_age,
                           studlab = new_study_id,
                           data = df_for_age_metan,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "age across studies",
                           subgroup = psy_or_med,
                           tau.common = FALSE)

@@

met_age_clin <- metagen(TE = mean_age,
                           seTE = se_age,
                           studlab = new_study_id,
                           data = df_for_age_metan_clin,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "age across clinical studies",
                           subgroup = psy_or_med,
                           tau.common = FALSE)

met_age_clin

# Now cbt studies only

df_for_age_metan_cbt <- df_for_age_metan %>%
  filter(new_study_id %in% cbt_meds_study_ids) %>% 
  filter(!(arm_effect_size == "cohens_d_active" & psy_or_med == 1 & !(type == "cbt")))

met_age_cbt <- metagen(TE = mean_age,
                           seTE = se_age,
                           studlab = new_study_id,
                           data = df_for_age_metan_cbt,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "age across cbt + med studies",
                           subgroup = psy_or_med,
                           tau.common = FALSE)

met_age_cbt

# Now look at baseline severity for overall psy vs med
# Need to calculate the variables first

df_for_baseline_severity <- df_long_for_metan %>% 
  group_by(new_study_id) %>%
  mutate(perc_baseline_overall = sum(perc_baseline_mean * baseline_n) / sum(baseline_n)) %>% 
  mutate(pooled_perc_sd_baseline = sqrt(sum((perc_baseline_sd^2) * (baseline_n - 1)) / sum(baseline_n - 1))) %>% 
  mutate(perc_se_baseline = pooled_perc_sd_baseline / sqrt(n_overall))

df_for_baseline_severity <- df_for_baseline_severity %>% 
  distinct(new_study_id, .keep_all = TRUE)

met_sev <- metagen(TE = perc_baseline_overall,
                           seTE = perc_se_baseline,
                           studlab = new_study_id,
                           data = df_for_baseline_severity,
                           sm = "",
                           fixed = FALSE,
                           random = TRUE,
                           method.tau = "REML",
                           hakn = TRUE,
                           title = "baseline severity",
                           subgroup = psy_or_med,
                           tau.common = FALSE)
